---
title: "g6R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{g6R}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(g6R)
library(shiny)
```

## Graph data

### Data formats

To setup a `{g6R}` graph, you first have to define __node__, __edges__ and __combos__ (collection of nodes), like so:

```{r}
nodes <- data.frame(id = 1:10)
nodes$label <- nodes$id
```

You could also pass these elements as __list__, which will be faster than the dataframe approach since under the hood, `{g6R}` has to send list to JavaScript:

```r
nodes <- lapply(seq_len(10), function(i) {
  list(id = i, label = i)
})
```

We then define some random edges:

```{r}
# Set a seed for reproducibility
set.seed(123)

# Define the number of edges to create (e.g., 200 random connections)
num_edges <- 5

# Generate random edges
edges <- data.frame(
  source = sample(nodes$id, num_edges, replace = TRUE),
  target = sample(nodes$id, num_edges, replace = TRUE)
)

edges$id <- paste0(edges$source, edges$target)
duplicated_id <- which(duplicated(edges$id) == TRUE)
if (length(duplicated_id)) {
  edges <- edges[-duplicated_id, ]
}
```

### Performance

If you are dealing with large graphs, you might want to use the `jsonUrl` parameter to __fetch__ the graph data from a hosted JSON file instead. If so, then leave `nodes`, `edges` and `combos` `NULL`. This allows to bypass the serialization of the graph data to JavaScript, which can be slow for large graphs. That's done in the `shinyAppDir(system.file("examples", "json", package = "g6R"))` example.

```{r g6R-json, eval=FALSE, echo = TRUE}
shinyAppDir(system.file("examples", "json", package = "g6R"))
```

```{r shinylive_url_2, echo = FALSE, results = 'asis'}
# extract the code from knitr code chunks by ID
code <- paste0(
  c(
    "webr::install(\"g6R\", repos = \"https://cynkra.github.io/blockr.webR/\")",
    knitr::knit_code$get("g6R-json")
  ),
  collapse = "\n"
)

url <- roxy.shinylive::create_shinylive_url(code, header = FALSE)
```

```{r shinylive_iframe_2, echo = FALSE, eval = TRUE}
tags$iframe(
  class = "border border-5 rounded shadow-lg",
  src = url,
  style = "zoom: 0.75;",
  width = "100%",
  height = "1100px"
)
```

## Initialise the graph

We leverage `g6` to create an __instance__ of our network:

```{r}
g6(nodes, edges, width = 200, height = 200)
```

As you can see, the nodes don't render well yet. Let's add it some layout.

## Layout

`g6_layout()` allows to pass any supported [layout](https://g6.antv.antgroup.com/en/manual/layout/overview). For this example, we select the `d3_force_layout()`:

```{r}
g <- g6(nodes, edges) |>
  g6_layout(d3_force_layout())
g
```

That's better! We could go further by displaying nodes __label__.

## Tweak options

`g6_options()` is your to go tool when it comes to change the style of the graph element such as nodes. Properties are selected from the [documentation](https://g6.antv.antgroup.com/en/manual/element/node/build-in/base-node#main-graphic-style):

```{r}
g <- g |>
  g6_options(
    node = list(
      style = list(
        labelBackground = TRUE,
        labelBackgroundFill = '#FFB6C1',
        labelBackgroundRadius = 4,
        labelFontFamily = 'Arial',
        labelPadding = c(0, 4),
        labelText = JS(
          "(d) => {
              return d.id
            }"
        )
      )
    )
  )
g
```

## Plugins

Plugins allowing users to improve the user experience by adding graphical components to the canvas like __minimaps__ or __tooltips__.
We can pass them inside `g6_plugins` either as a character string with a reference to the plugin name or using the correponding function to pass more configuration options:

```r
# Use defaults
g6_plugins("minimap")

# Custom options
g6_plugins(
  minimap(size = c(100, 100))
)
```

```{r}
g <- g |>
  g6_plugins(
    minimap(size = c(100, 100))
  )
g
```

## Behaviors

Behaviors correspond to interactions between users and the graph elements, such as dragging the nodes and selecting nodes. Be mindful that not all behaviors are compatible: dragging canvas and canvas elements would require to specify different triggers for each behavior.
With `{g6R}` behaviors can be added with `g6_behaviors`, like plugins:

```{r}
g <- g |>
  g6_behaviors(
    "zoom-canvas",
    drag_element_force(fixed = TRUE),
    click_select(
      multiple = TRUE,
      onClick = JS(
        "(e) => {
            console.log(e);
          }"
      )
    ),
    brush_select()
  )
g
```

Notice that we can pass __callback__ functions from R to JavaScript. This is useful in combination with Shiny to set custom inputs, for instance.
